---
title: "Multiclass Diabetes Health Trend Analysis"
author: "Unnati Agrawal"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(nnet)
 
```{r load and clean data}
# Load dataset
df1 <- read_csv("data/diabetes_012_health_indicators_BRFSS2015.csv") %>% clean_names()

# Convert target to factor
df1$diabetes_012 <- as.factor(df1$diabetes_012)

# Preview structure
glimpse(df1)

```{r eda}
df1 %>%
  count(diabetes_012) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = diabetes_012, y = prop, fill = diabetes_012)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Diabetes Class Proportions (0 = No, 1 = Pre, 2 = Diabetes)", y = "Proportion") +
  theme_minimal()
 
ggplot(df1, aes(x = bmi, fill = diabetes_012)) +
  geom_histogram(position = "identity", bins = 30, alpha = 0.6) +
  labs(title = "BMI Distribution by Diabetes Status", x = "BMI", y = "Count") +
  theme_minimal()

```{r multinomial logistic regression}
# Fit multinomial logistic regression
model1 <- multinom(diabetes_012 ~ bmi + high_bp + gen_hlth + age + phys_activity, data = df1)

# Model summary
summary(model1)

# Compute z-values and p-values
z <- summary(model1)$coefficients / summary(model1)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
round(p, 5)

