---
title: "Binary Diabetes Health Trend Analysis"
author: "Unnati Agrawal"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)

## 📂 1. Load and Clean the Data

```{r load-and-clean}
# Load dataset from file
df <- read_csv("data/diabetes_binary_5050split_health_indicators_BRFSS2015.csv")

# Clean column names
df <- df %>% clean_names()

# Convert diabetes_binary to factor for plotting & modeling
df$diabetes_binary <- as.factor(df$diabetes_binary)

# View the structure
glimpse(df)


```{r exploratory data analysis}
df <- read_csv("outputs/cleaned_diabetes_data.csv")
df$diabetes_binary <- as.factor(df$diabetes_binary)


# BMI distribution
ggplot(df, aes(x = bmi, fill = diabetes_binary)) +
  geom_histogram(position = "identity", alpha = 0.6, bins = 30) +
  labs(title = "BMI Distribution by Diabetes Status") +
  theme_minimal()

ggsave("plots/bmi_by_diabetes.png")

# General health vs diabetes
df %>%
  group_by(gen_hlth, diabetes_binary) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = gen_hlth, y = count, fill = diabetes_binary)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "General Health by Diabetes Status", x = "General Health") +
  theme_minimal()

ggsave("plots/gen_health_by_diabetes.png")

```{r logistic regression model}
df <- read_csv("outputs/cleaned_diabetes_data.csv")

# Fit a logistic regression
model <- glm(diabetes_binary ~ bmi + high_bp + gen_hlth + age + phys_activity,
             data = df, family = binomial)

# Summary of the model
summary(model)
